---
title: "cryptography"
author: "Max Lang"
date: "10/27/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(checkmate)
```

# find_palindrome
This function returns TRUE if a given string is a palindrome, FALSE otherwise.

## Input
* `input`: A `character(1)`.

## Output
* Logical vector of length one

## Code
```{r}
is_palindrome <- function(input) {
  assertString(input)
  str <- strsplit(toupper(input), "")[[1]]
  str <- str[str %in% LETTERS]
  all(str == rev(str))
}
```
## Worked example
```{r}
is.a.palindrome <- "Was it a car or a cat I saw?"
not.a.palindrom <- "Hello, what is your name?"
is_palindrome(is.a.palindrome)
is_palindrome(not.a.palindrom)
```

# caesar_cipher
This function can encrypt and decrypt plain text using the caesar cyper.

# Input
* plaintext: Text to encrypt / decrypt
* key: Letter used as the key.
* decrypt: Set true if you want to decrypt encrypted text

# Output
Encrypted/decrypted text

## Code
```{r}
caesar_cipher <- function(plaintext, key, decrypt = FALSE) {
  assertString(plaintext, pattern = "^[A-Z ]*$")
  assertString(key, pattern = "^[A-Z ]$")
  assertLogical(decrypt)

  code <- c(" ", LETTERS)
  plaintext <- strsplit(plaintext, "")[[1]]
  decryptmultiplier <- if (decrypt) -1 else 1

  numbers <- (match(plaintext, code) - 1) + (match(key, code) - 1) * decryptmultiplier
  numbers <- numbers %% 27
  paste(code[numbers + 1], collapse = "")
}

```

## Worked example (encrypt)
```{r}
encrypted.text <- caesar_cipher("HELLO THIS IS A VERY SECRET MESSAGE", 
                                key = "L", 
                                decrypt = FALSE)

encrypted.text
```

## Worked example (decrypt)
```{r}
caesar_cipher(encrypted.text, key = "L", decrypt = TRUE)
```

# Breaking the caesar cypher
We will use letter frequencies based on this Wiki article <https://en.wikipedia.org/wiki/Letter_frequency>.
```{r}
letterfrequencies <- 1 / 100 * c(
  A = 6.756, B = 1.234, C = 2.302, D = 3.518, E = 10.508, F = 1.843, G = 1.667,
  H = 5.041, I = 5.763, J = 0.127, K = 0.639, L = 3.330, M = 1.990, N = 5.583,
  O = 6.210, P = 1.596, Q = 0.079, R = 4.953, S = 5.234, T = 7.492, U = 2.282,
  V = 0.809, W = 1.952, X = 0.124, Y = 1.633, Z = 0.061, ` ` = 17.272)
```

# text_log_likelihood
This function estimates the log-likelihood that a given text is, in fact, a non-encrypted plain text, using the distribution of letters.

## Input
- text: A `character(1)` string made up of upper case letters and space

## Output
A scalar `numeric` giving the log likelihood of a given text.

## Code 
```{r}
text_log_likelihood <- function(text) {
  # your code
  assertString(text, pattern = "^[A-Z ]+$")
  sum(log(letterfrequencies[strsplit(text, "")[[1]]]))
}
```
## Worked Example
```{r}
text_log_likelihood("HI WHAT IS UP")
```

# estimate_key
This function estimates the most likely key for a given ciphertext.
This is the key that generates a text that is most likely according to `text_log_likelihood`. The possible keys are the 26 letters as well as the space (`" "` -- this one does not change the text).

## Input
- `ciphertext`: A `character(1)` string made up of upper case letters and space

## Output: a list with two entries:
 - `key`: `character(1)` giving an upper case letter or space.
 - `log.likelihood`: `numeric(1)` giving the log likelihood of the text when decrypting with this key.
 
## Code
```{r}
estimate_key <- function(ciphertext) {
  assertString(ciphertext, pattern = "^[A-Z ]+$")
  keys <- c(LETTERS, " ")
  result <- vapply(keys, function(k) {
    text_log_likelihood(caesar_cipher(ciphertext, k, TRUE))
  }, numeric(1))
  list(key = names(which.max(result)), log.likelihood = max(result))
}
```
## Worked example
```{r}
cipher <- caesar_cipher("HELLO THIS IS A SECRET MESSAGE TO CAESAR", key = "B", decrypt = FALSE)
estimate_key(cipher)
```

# break_caesar
This function uses the previous function to break the caesar cipher.

## Input:
* A character vector of length one containing an (caesar) encrypted text with unknown key.

## Output:
Hopefully the correct message.

## Code
```{r}
break_caesar <- function(cipher) {
  assertCharacter(cipher, any.missing = FALSE, min.len = 1)
  infos <- do.call(rbind, lapply(cipher, function(x) as.data.frame(estimate_key(x))))
  needle <- which.max(infos$log.likelihood)
  caesar_cipher(cipher[[needle]], infos$key[[needle]], decrypt = TRUE)
}
```
## Worked example
```{r, include=FALSE}
ciphertext <- caesar_cipher(toupper("Meanwhile with the legion which he had with him and the soldiers which had assembled from the Province he carries along for nineteen Roman miles a wall to the height of sixteen feet and a trench from the Lake of Geneva which flows into the river Rhone to Mount Jura which separates the territories of the Sequani from those of the Helvetii When that work was finished he distributes garrisons and closely fortifies redoubts in order that he may the more easily intercept them if they should attempt to cross over against his will When the day which he had appointed with the embassadors came and they returned to him he says that he can not consistently with the custom and precedent of the Roman people grant any one a passage through the Province and he gives them to understand that if they should attempt to use violence he would oppose them The Helvetii disappointed in this hope tried if they could force a passage some by means of a bridge of boats and numerous rafts constructed for the purpose others by the fords of the Rhone where the depth of the river was least sometimes by day but more frequently by night but being kept at bay by the strength of our works and by the concourse of the soldiers and by the missiles they desisted from this attempt"), "P")
```
```{r}
ciphertext
break_caesar(ciphertext) # The passage is from The Gallic Wars by Julius Caesar
```


