---
title: "regular-expression"
author: "Max Lang"
date: "10/27/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(checkmate)
```
# `find_url``
This function will find URLs in long text and splits them into the respective components
such as protocol, domainname etc.

#ä Input
* `input`: A `character(1)` containing the input string.

## Output:
* A data.frame with columns protocol, domainname and path.

## Code
```{r}
find_url <- function(input) {
  assertString(input)
  rx <- "(?<![[:alpha:]])(https?)://([[:alnum:]][-[:alnum:]]*(?:\\.[[:alnum:]][-[:alnum:]]*)+)/([-a-z[:digit:]/]*)"

  matches <- gregexpr(rx, input, perl = TRUE, ignore.case = TRUE)
  urls <- regmatches(input, matches)[[1]]

  m <- regexec(rx, urls, perl = TRUE, ignore.case = TRUE)
  urlparts <- regmatches(urls, m)
  res <- t(vapply(urlparts, identity, character(4), USE.NAMES = FALSE))
  res <- res[, -1]
  res <- as.data.frame(res, stringsAsFactors = FALSE)
  colnames(res) <- c("protocol", "domainname", "path")
  res
}
```
## Worked example
```{r}
string <- "https://www.google.com/ is probably the most popular url, sometimes used ashttp://www.google.com/ if one forgets to use the more secure https:// protocol. Another URL that is popular is http://youtube.com which is owned by the same company. On some computers, it is possible to find a website at http://localhost/ if the computer runs a webserver. This file can be found at HTTPS://GITHUB.COM/MAXMLANG/STRUCTURED-PROGRAMMING."
find_url(string)
```

# ---------------------------------------------------------------------------------------------------------------
# Zweite Hausarbeit
# Programmieren mit Statistischer Software
# SS 2022
# ---------------------------------------------------------------------------------------------------------------
# Before you start: Please register for this exam on moodle and make sure you have read *and signed* the angabe.pdf
# included in this repository.
# For this exercise, only base R packages and checkmate are allowed!
# ---------------------------------------------------------------------------------------------------------------

# In this exercise you are provided with a rds file storing a vector of some made up file names.
# To ensure everything works correctly, it's better to just not touch the rds file, other than reading it into R
# like this:
exFileNames <- readRDS("assets/filenames.rds")
# You can look at `exFileNames` in the RStudio environment pane or run e.g. `head(exFileNames)` in the console
# to see the first few files (or rather file names).
# You're going to work with that vector of file names throughout this exercise.
# Imagine it contains the result of running `list.files()` on your local computer,
# showing you all the files in e.g. the directory you keep your study projects in.
#
# Luckily the files follow a more or less consistent format. It seems like Jenny Bryan's advice on naming things was
# followed at least partially (see https://speakerdeck.com/jennybc/how-to-name-files).
#
# FORMAT: File names have components "date", "project", "revision" and "extension", always in that order.
# - File names start with the date they were created, in year-month-date order, but sadly separators "-" and "_" were
#   used inconsistently, and sometimes no separator was used at all: It could either be 2022-02-23 or 2022_02_23
#   or 20220223 for the same date. At least they always start with a 4-digit year, and month and day components
#   are always two digits as well: It's 20220109 rather than 202219 for example.
# - Project names could be any alphanumeric string it seems. They are separated from the date via "-".
# - The revision follows the project name, and the two are always separated by "_", which seems good to know.
# - The file extensions come at the end of the string, after the "." as usual. There are multiple file extensions
#   with inconsistent capitalization, for example both "CSV" and "csv" of course refer to the same type of file.
#   Also, some extensions are three characters, and others such as for R scripts are shorter.
#   The files we care about all have extensions consisting of letters only, there are no .mp3 files for example.
#
# For evaluation purposes, a different randomly generated set of file names will be used, which might consist
# of only one file name, or 50, or 100. They will have the same format but you should not rely on the exact contents
# of the example file names we have provided with `exFileNames`.

# Exercise 01: File Counts ---------------------------------------------------------------------------------------

# With the list of files at hand, you first wonder what the most common file types are as identified by their extension.
# You also wonder how many files there are per year.
# Write a function that takes the vector of file names and returns a named integer vector with counts by either year
# or file type (similar to what table() produces, see note below).
#
# Function arguments:
# - `filenames` (no default): A `character` vector containing one or more file names.
#                             Look at `exFileNames` for example input.
# - `by`: A `character` with allowed values `"year"` or `"type"`, defaulting to `"type"`, that specifies whether to
#         count files by year or by type (= file extension).
#
# Your function should use checkmate assertions to ensure that the input file names are of the expected format, i.e.
# starting with a 4-digit year and having a file extension at least. The other parts of the file names are not
# relevant to this particular task, so you can choose to ignore them.
# Missing or empty (`character(0)`) values are not allowed.
#
# Return value:
# A named `integer` vector, sorted in ascending frequency of counts (lowest to highest from left to right) and
# names corresponding to respective year or file type. File extensions should be converted to lower case before
# counting to avoid duplication due to different capitalizations. The names of the output vector should therefore
# also be lower case for `by = "type"`.
#
# EXAMPLE for valid input from exFileNames:
# exampleFiles <- c("2019-08-27-Assignment2_beta.r", "20200422-fermat-proof_DONE.Rds",
#                   "2020_02_19-Assignment2_pls-help.Rds")
# ex01FileCount(exampleFiles, by = "type")
# Output: c("r" = 1L, "rds" = 2L)
#
# ex01FileCount(exampleFiles) -> Same output, since `by = "type"` should be the default.
#
# ex01FileCount(c("20190313-sample_final.Rds", "20210806-simulation-data_final.RDS"), by = "type")
# Output: c("rds" = 2L)
#
# ex01FileCount(exampleFiles, by = "year")
# Output: c("2019" = 1L, "2020" = 2L)
#
# Please note that the example output will look like this in the console:
# 2019 2020
#    1    2
# And that the example output above uses e.g. `2L` to make clear that the numbers are supposed to be integers
# rather than just numerics.
#
# EXAMPLE for invalid input:
# exampleFiles <- c("Assignment2_beta.r", "20200422-fermat-proof_DONE.Rds", "2020_02_19-Assignment2_pls-help.Rds")
# ex01FileCount(exampleFiles) -> Checkmate assertion error since one file does not start with a year/date as expected.
#
# exampleFiles <- c("2019-08-27-Assignment2_beta.r", "9999tempfile")
# ex01FileCount(exampleFiles) -> Checkmate assertion error since one file does not have a file extension.
#
# ex01FileCount("") -> Checkmate assertion error since input file is an empty string.
# ex01FileCount(list(file = "2022-03-13-myFile.rds")) -> Checkmate assertion error, input is not a character vector.
#
# ex01FileCount(c("20190313-sample_final.Rds", "20210806-simulation-data_final.RDS"), by = "extension")
# -> Checkmate assertion error since `by` should be either `"year"` or `"type"`, nothing else
#
# Note: If you use table(), wrap it in c(), e.g. c(table(...)) to make the return value
# a named vector and not an object of class "table". This part is similar to an earlier homework.
# You can check your result with `is.integer()` (should be `TRUE`), and `class()` (should be integer), while
# `names()` should give you a character vector of e.g. years for `by = "year"`.
ex01FileCount <- function(filenames, by = "type") {
  # Check only for year/file extension components this time, but ensure no missings or empty values.
  assertCharacter(filenames, pattern = "^\\d{4}.*\\.([a-zA-Z]+)$", min.chars = 1, any.missing = FALSE, min.len = 1)
  assertChoice(by, choices = c("year", "type"))

  # Avoid if-else block by switching patterns for each case of "by", makes it easier to extend to other `by`'s
  typePattern <- switch(by, year = "^\\d{4}", type = "([a-zA-Z]+)$")
  # extract sub-matches for year/type, and iterate over them with vapply() to extract the list-element
  # since regmatches yields a list.
  fileSplit <- vapply(regmatches(filenames, regexec(typePattern, text = filenames)), function(x) x[[1]], "")

  # 1. lower case the result
  # 2. count occurrences with table, make it a named integer vector with c() (rather than class "table")
  # 3. sort it in ascending order
  # Stuff like this is why people like using pipes.
  sort(c(table(tolower(fileSplit))))
}

# Exercise 02: File Names to DF -----------------------------------------------------------------------------------

# Now that you can give yourself a broad overview of the files, you want to represent them in a more useful format
# like a data.frame.
# Write a function that takes a vector of file names and returns a `data.frame` with one variable per component.
##
# Function input:
# - `filenames` (no default): A `character` vector containing one or more file names.
#                             Look at `exFileNames` for example input.
#
# The input should be validated regarding the naming format described above, where invalid file names should
# lead to a checkmate assertion error, as well as NAs or empty (`character(0)`) values.
# Note that in this function you need all parts of the file name rather than just the year and extension like in
# ex01FileCount(), so your validation pattern should reflect that.
#
# Return value:
# A `data.frame` with the following `character` variables in that order(!):
# - `date`: The date in ISO 8601 format, yyyy-mm-dd, e.g. "2022-06-27". Only use "-" to separate date components.
# - `project`: The project part of the file name, as described above, e.g. "simulation-data".
# - `revision`: The revision part of the file name, as described above, e.g. "beta" or "done"
# - `ext`: The file extension, without ".", e.g. "rds"
#
# All variables should be of type `character` and lower case to avoid duplication due to differences in capitalization.
#
# EXAMPLES for valid input:
# exampleFiles <- c("2020_06_20-sample_DONE.Csv", "2021-06-19-exam-prep_DONE.Rds", "20211224-homework_final2.pdf")
# Return value:
# data.frame(
#   date = c("2020-06-20", "2021-06-19", "2021-12-24"),
#   project = c("sample", "exam-prep", "homework"),
#   revision = c("done", "done", "final2"),
#   ext = c("csv", "rds", "pdf")
# )
#
# You can also use str() to check the structure of your function's return value, which should look like this:
# > str(ex02FileNamesDF(exampleFiles))
# 'data.frame': 3 obs. of  4 variables:
# $ date    : chr  "2020-06-20" "2021-06-19" "2021-12-24"
# $ project : chr  "sample" "exam-prep" "homework"
# $ revision: chr  "done" "done" "final2"
# $ ext     : chr  "csv" "rds" "pdf"
#
# And to avoid confusion regarding row names or other attributes, you can check them with attributes():
# > attributes(ex02FileNamesDF(exampleFiles))
# $names
# [1] "date" "project""revision" "ext"
#
# $row.names
# [1] 1 2 3
#
# $class
# [1] "data.frame"
#
# If you normally construct the `data.frame` it will automatically have row names starting with 1 up to the number of
# rows, and there will be no other classes or attributes attached (this is what we expect / want).
# For safety we still encourage you to check that your function behaves as you expect!
#
# exampleFiles <- "2020_06_20-sample_DONE.Csv"
# Return value:
# data.frame(
#   date = c("2020-06-20"),
#   project = c("sample"),
#   revision = c("done"),
#   ext = c("csv")
# )
#
# EXAMPLES for invalid input:
# (All of these should again lead to a checkmate assertion error!)
# ex02FileNamesDF("220515-thesis_final.Rds") -> Date format not correct
# ex02FileNamesDF(list(filename = "2020-11-03-sample_final2.pdf")) -> Not a character vector
# ex02FileNamesDF("2022-04-08.rds") -> Does not match format (no project or revision component)
# ex02FileNamesDF("2020_01_15-results.rds") -> Does not contain a revision component
#
# NB: If these file names seem irritating to you: You're right. Real examples of bad file names are much worse!
ex02FileNamesDF <- function(filenames) {
  # capture groups for year, month, date, project, revision, extension
  pattern <- "^(\\d{4})[-_]?(\\d\\d)[-_]?(\\d\\d)-([[:alnum:]-]*)_([[:alnum:]-]*)\\.([a-zA-Z]+)$"

  # Check pattern, and ensure no missing/empty values etc.
  # where assertString in ex01AssertAddress insists on a length-1 input, assertCharacter allows other lengths
  # and we ensure we get at least a length-1 vector.
  assertCharacter(filenames, any.missing = FALSE, pattern = pattern, min.chars = 1, min.len = 1)

  # Extract matches from capture groups
  submatches <- regmatches(filenames, regexec(pattern = pattern, text = filenames))
  # Iterate over matches (a list containing submatches)
  matchlistdf <- lapply(submatches, function(x) {
    # Convert to lower case and exclude first element (the full match, we don't need it)
    x <- tolower(x[-1])
    # Name vector according to specification, order is ensured to be the same as input anyway.
    names(x) <- c("year", "month", "day", "project", "revision", "ext")
    # Create data.frame (1 row) in format as required
    data.frame(
      # Paste the date together from its components, separated by "-".
      date = paste(x[["year"]], x[["month"]], x[["day"]], sep = "-"),
      project = x[["project"]], revision = x[["revision"]], ext = x[["ext"]]
    )
  })

  # matchlistdf is a list containing 1-row data.frames, we can successively bind them row-wise with do.call(rbind, ...)
  # The lapply-do.call-rbind pattern is mostly equivalent to purrr::map_dfr(), so it's very handy if you don't want
  # to depend on purrr.
  do.call(rbind, matchlistdf)
}
