---
title: "basic-vector-ops"
author: "@MaxMLang (Github)"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(checkmate)
```

# `bin_count`
This function counts the number of observations in a specific bin. The user
can specify the cutoff points. The function returns a named vector. 

## Inputs
* `observations`: A numeric vector containing observations
* `cuts`: A numeric vector with cutoff points
* `binnames`: A character vector specifying the labels for the bins in the output

## Output
* A named vector with the number of observations in each bin.

## Code
```{r}
bin_count <- function(observations, cuts, binnames) {
  cumobs <- vapply(c(cuts, Inf), function(x) sum(observations < x), 0)
  result <- diff(c(0, cumobs))
  names(result) <- binnames
  result
}
```
## Worked example
```{r}
observations <-  c(18, 22, 28)
cuts  <- c(15, 19.5, 24)
binnames <-  c("Severely underweight", "Underweight", "Healthy", "Overweight")
bin_count(observations, cuts, binnames)
```
# `bin_data`
This function bins each data point into a category / bin. The user
can specify the cutoff points. The function returns a named vector. 

## Inputs
* `observations`: A numeric vector containing observations
* `cuts`: A numeric vector with cutoff points
* `binnames`: A character vector specifying the labels for the bins in the output

## Output
* A named vector with length of the input data that is now binned.
```{r}
bin_data <- function(observations, cuts, binnames) {
  cut(observations, breaks = c(-Inf, cuts, Inf), labels = binnames, right = FALSE, ordered_result = TRUE)
}
```
## Worked Example
```{r}
observations <- c(20, 23, 28)
cuts <- c(16, 18.5, 25)
binnames <-  c("Severely underweight", "Underweight", "Healthy", "Overweight")
bin_data(observations, cuts, binnames)
```

# `mark_divisible`
This function marks every number that is divisible by the specified the divisor 
up to the specified length of the sequence.

## Inputs
* `up.to`: A numeric vector specifying the end of the sequence
* `divisor`: A numeric vector specifying the divisor
* `marker`: A character vector specifying the marker

## Output
* A named vector with length of the input data that is now binned.
```{r}
mark_divisible <- function(up.to, divisor = 2, marker = "even") {
  # your code
  sequence <- seq_len(up.to)
  ret <- as.character(sequence)
  is.divisible <- sequence %% divisor == 0
  ret[is.divisible] <- marker
  ret
}
```
## Worked Example
```{r}
up.to <- 50
divisor <- 9
marker <-  "X"
mark_divisible(up.to, divisor, marker)
```

# `find_fibonacci`
Finds the n th Fibonacci number. It is a recursive approach.
## Inputs
* `n` a non-negative scalar integer value

## Output
* scalar value: the n'th Fibonacci number.

```{r}
find_fibonacci <- function(n) {
  assertCount(n, tol = 1e-100)
  if (n <= 1) {
    return(n)
  }
  find_fibonacci(n - 1) + find_fibonacci(n - 2)
}
```
## Worked example
```{r}
find_fibonacci(5)
find_fibonacci(14)
```
# `vec_threshold`
Returns all elements of the vector that are at least onee number greater than the
threshold.

## Inputs
* `vectors`: a list of `numeric` vectors
* `threshold`: a scalar numeric

## Outputs
Character vector containing all elements from `vectors` that contain at least one number greater than `threshold`.
#
# Examples:
# ex03VectorThreshold(
#   list(numeric(0), 1:3, 8:11, c(-100, 100)),
#   threshold = 10)
# --> list(8:11, c(-100, 100))
# ex03VectorThreshold(list(), 10) --> list()
# ex03VectorThreshold(list(numeric(0), 0, -1), 10) --> list()
# ex03VectorThreshold(10, 10) --> ERROR (`vectors` is not a list of numerics)
# ex03VectorThreshold(list(10), "10") --> ERROR (threshold is not a numeric)
#
# You may want to use `Filter`, and the solution of ex02VectorCondition may be useful here.
```{r}
vec_threshold <- function(vectors, threshold) {
  assertList(vectors, any.missing = FALSE)
  Filter(vec_condition(threshold), vectors)
}

# helper function
vec_condition <- function(threshold) {
  assertNumber(threshold)
  function(vect) {
    assertNumeric(vect, any.missing = FALSE)
    any(vect > threshold)
  }
}
```
## Worked example
```{r}
vec_threshold(list(numeric(0), 1:3, 8:11, c(-100, 100)), threshold = 10)
vec_threshold(list(numeric(0), 0, -1), 10)
```
