---
title: "basic-matrix-df-ops"
author: "@MaxMLang (Github)"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(checkmate)
```

# `chess_black_sum`
This functions sums all numbers that would be on black fields on a chessboard, here
represented as a matrix.

## Inputs
* matrix: A matrix representing a chessboard

## Output
* A numeric vector containing the number of black fields

## Code
```{r}
chess_black_sum <- function(matrix) {
  sum(matrix[(row(matrix) + col(matrix)) %% 2 == 1])
}
```

## Worked example
```{r}
matrix <- matrix(data= 1:9, nrow = 3)
matrix
chess_black_sum(matrix)
```

# `which_mat_min`
This function finds the indexed positon of the minimum value in the matrix.

## Inputs
* matrix: A matrix 

## Output
* A numeric vector containing the minimum indices i,j

## Code
```{r}
which_mat_min <- function(matrix) {
  # your code
  wm <- which.min(matrix)
  as.numeric(c(row(matrix)[wm], col(matrix)[wm]))
}
```

## Worked example
```{r}
matrix <- matrix(sample(100, size = 25), ncol = 5)
matrix
which_mat_min(matrix)
```

## `diag_df`
This function returns the diagonal of a dataframe.

## Inputs
* `df`: a data.frame with arbitrary rows and columns

## Output
* unnamed `list` of the diagonal elements of `df`

## Code
```{r}
diag_df <- function(df) {
  # your code
  assertDataFrame(df)
  lapply(seq_len(min(nrow(df), ncol(df))), function(i) df[i, i])
}
```
## Worked example
```{r}
df <- data.frame(v1= sample(50, 10), v2= sample(50, 10), v3= sample(50, 10))
df
diag_df(df)
```

