---
title: "control-structures"
author: "@MaxMLang (Github)"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# `find_opposite`
Finds the 'opposite' of a value in a two column dataframe.

## Inputs
* database: A two column dataframe (colnames(value, opposite)), where each row represents a pair
* value: A value in the dataframe you want to find the opposite of.

## Output
* The respective opposite value
```{r}
find_opposite <- function(database, value) {
  if (value %in% database$value) {
    position <- which(database$value == value)[[1]]
    database$opposite[[position]]
  } else {
    position <- which(database$opposite == value)[[1]]
    database$value[[position]]
  }
}
```
## Worked example
```{r}
data <- data.frame(list(value = c("positive", "white", "Ying", "left"),
                opposite = c("negative", "black", "Yang", "right")))

find_opposite(data, "positive")
find_opposite(data, "Ying")
# works both ways
find_opposite(data, "right")
```

# `cellular_auto`
A "cellular automaton" is a discrete model of state evolution. We consider a state
of a vector with "cells" of values 0 or 1, for example the vector c(0, 1, 0, 1, 0, 0, 0).
The state now changes according to a special rule: Every cell changes to state 1
whenever its immediate left neighbour has state 1. A cell that has state 1 remains at state 1.
Each cell that has no left neighbour or a left neighbour that is 0, and that is also 0 itself, remains 0 otherwise.

## Inputs
* start.state: A vector only containing 0 and 1 representing he initial state of the cell.

## Output
A list of n process steps until the process is completed. 

## Code
```{r}
cellular_auto <- function(start.state) {
  results <- list(start.state)
  state <- start.state
  repeat {
    # shift the state to the right by 1
    state.shifted <- c(0, head(state, -1))
    # every state where state or state.shifted is 1 should be 1, others 0
    state.next <- ifelse(state == 1 | state.shifted == 1, 1, 0)
    # if nothing changed, we are done
    if (all(state.next == state)) break
    results[[length(results) + 1]] <- state <- state.next
  }
  results
}
```
## Worked example
```{r}
startcell <- c(0,0,1,0,1,0,0,1,0)
cellular_auto(startcell)
```



