---
title: "testing"
author: "Max Lang"
date: "10/29/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(sna)
library(network)
library(checkmate)
```

# is_connected
This function checks if two nodes in an undirected Graph are connected. The graph of n vertices can be represented by a symmetric n x n matrix X of values 0 and 1, where the element in row a and column b, i.e. X[a, b] is 1 if the edges a and b are connected by an edge, and 0 otherwise. (The matrix is symmetric, so X[a, b] == X[b, a], since a connection from a to b implies a connection from b to a.). This is the so-called "Adjacency matrix".


```{r, include= FALSE}
# random graph
set.seed(2)
net <-  rgraph(4, mode = "graph", tprob = 0.5)
net <-  network(net, directed = FALSE, vertex.attrnames = LETTERS[1:4])


plot(net)
```

For the graph above this matrix could for example look like this if we numerate from the left clockwise, then the matrice would look like the following.

$$
A = 
\begin{pmatrix}
0 & 1 & 0 & 0 \\
1 & 0 & 1 & 1 \\
0 & 1 & 0 & 1 \\
0 & 1 & 1 & 0
\end{pmatrix}
$$

So this function takes an adjacency matrix as input `adjacency`, an integer value `beginning` and an integer value `end`, and checks whether there is a series of edges that connect the `beginning` to the `end` vertex, returning
TRUE if it is the case and FALSE if not. 

# Input
adjacency: An adjacency matrix 
beginning: An integer value
end: An integer value

# Output
A logical vector returning `TRUE` if it is the case.

# Code
```{r}
is_connected <- function(adjacency, beginning, end) {
  assertMatrix(adjacency, any.missing = FALSE, mode = "numeric")
  if (nrow(adjacency) != ncol(adjacency)) {
    stop("adjacency must be a square matrix")
  }
  assertInt(beginning, lower = 1, upper = nrow(adjacency))
  assertInt(end, lower = 1, upper = nrow(adjacency))
  if (!all(adjacency == t(adjacency))) {
    stop("matrix is not symmetric")
  }
  visited <- numeric(0)
  current.nodes <- beginning

  # if we hit a node we've seen before, then we quit and give up
  while (length(current.nodes)) {
    if (end %in% current.nodes) {
      # we can reach the 'end' from here
      # --> beginning and end are connected
      return(TRUE)
    }
    visited <- c(visited, current.nodes)
    current.nodes <- which(apply(adjacency[current.nodes, , drop = FALSE], 2, sum) > 0)
    # choose a node that was not visited before
    current.nodes <- setdiff(current.nodes, visited)
  }
  FALSE
}
```
# Worked example
So for the graph above we can check with the following code.
```{r}
is_connected(matrix(c(0,1,0,0,1,0,1,1,0,1,0,1,0,1,1,0), nrow = 4), 1, 4)
```

